---
title: "Final Project Proposal"
format: pdf
author: "Diego Moylan"
---

```{r}
#| include: false
library(tidyverse)
```


Team: Diego Moylan

The data I plan on working with comes from Kaggle, specifically these links:

https://www.kaggle.com/datasets/thedevastator/jobs-dataset-from-glassdoor?select=eda_data.csv (2017)

https://www.kaggle.com/datasets/andrewmvd/data-analyst-jobs (June 2020)

https://www.kaggle.com/datasets/louise2001/us-cities 

They represent data science and related jobs scraped from Glassdoor (and possibly Indeed) with salary and location information. The US cities dataset is included almost entirely because it has latitude/longitude information that I can join onto the location of the jobs. This should allow me to make some neat looking maps, which would look good on a dashboard. So far my best idea is to graph the jobs on a map, with color and size (of a dot on a city) representing salary and number of jobs (or vice versa).


```{r}
#| output: false
cities = read_csv(
  "uscities.csv",
  show_col_types = FALSE,
  col_select = c("city", "state_id", "lat", "long" = "lng", "population")
)
```

```{r}
#| output: false
jobs1 = read_csv(
  "eda_data.csv",
  show_col_types = FALSE,
  col_select = c("title" = "Job Title", "salary_est" = "Salary Estimate",
                 "company" = "Company Name", "location" = "Location",
                 "size" = "Size"
                 )
)
```

```{r}
#| output: false
jobs2 = read_csv(
  "DataAnalyst.csv",
  show_col_types = FALSE,
  col_select = c("title" = "Job Title", "salary_est" = "Salary Estimate",
                 "company" = "Company Name", "location" = "Location",
                 "size" = "Size"
                 )
)
```

```{r}
#| ouput: false
jobs_join = dplyr::union(jobs1, jobs2)
```


```{r}
#| ouput: false
#| warning: false
jobs_clean <- jobs_join |>
  separate_wider_delim(
    location, 
    names = c("city", "state_id"), 
    delim = ",",
    too_few = ("align_start"),
    too_many = ("drop")
    ) |>
  mutate(salary_est = (str_remove_all(salary_est, "[^0-9-K]"))) |>
  separate_wider_delim(
    salary_est,
    names = c("min", "max"),
    delim = "-",
    too_few = ("align_start"),
    too_many = ("drop")
  ) |>
  mutate(
    
    #no K means it is an hourly wage, which can closely approximate
    #annual wages by multiplying by 2000x. Probably better than NA
    
    min = case_when(
      is.na(min) ~ NA,
      str_detect(min, "K") ~ 1000*as.integer(str_remove_all(min, "[^0-9]")),
      .default = 2000*as.integer(min)
    ),
    max = case_when(
      is.na(max) ~ NA,
      str_detect(max, "K") ~ 1000*as.integer(str_remove_all(max, "[^0-9]")),
      .default = 2000*as.integer(max)
    ),
    avg = as.integer((min + max) / 2)
  ) |>
  relocate(avg, .after = max)

```


```{r}
#| output: false
#NA values for all lat/long/pop from cities
#no idea why, but that is an issue for later
jobs <- jobs_clean |>
  left_join(cities, join_by(city, state_id))
```

```{r}
jobs |>
  summarize(
    avg_salary_us = mean(avg, na.rm = TRUE)
  )
```
```{r}
jobs |> 
  count(city, name = "jobs") |>
  arrange(desc(jobs))
```










