---
title: "Working Draft"
format: pdf
author: "Diego Moylan"
---

```{r}
#| include: false
library(tidyverse)
library(sf)
library(ggspatial)
library(rnaturalearth)
library(rnaturalearthdata)
library(leaflet)

```


Team: Diego Moylan

The data I plan on working with comes from Kaggle, specifically these links:

https://www.kaggle.com/datasets/thedevastator/jobs-dataset-from-glassdoor?select=eda_data.csv (2017)

https://www.kaggle.com/datasets/andrewmvd/data-analyst-jobs (June 2020)

https://www.kaggle.com/datasets/louise2001/us-cities 

They represent data science and related jobs scraped from Glassdoor (and possibly Indeed) with salary and location information. The US cities dataset is included almost entirely because it has latitude/longitude information that I can join onto the location of the jobs. This should allow me to make some neat looking maps, which would look good on a dashboard. So far my best idea is to graph the jobs on a map, with color and size (of a dot on a city) representing salary and number of jobs (or vice versa).




```{r}
#read in data

#| warning: false

cities = read_csv(
  "./Data/uscities.csv",
  show_col_types = FALSE,
  col_select = c("city", "state_id", "lat", "long" = "lng", "population")
)

jobs_1 = read_csv(
  "./Data/eda_data.csv",
  show_col_types = FALSE,
  col_select = c("title" = "Job Title", "salary_est" = "Salary Estimate",
                 "company" = "Company Name", "location" = "Location",
                 "size" = "Size"
                 )
)

jobs_2 = read_csv(
  "./Data/DataAnalyst.csv",
  show_col_types = FALSE,
  col_select = c("title" = "Job Title", "salary_est" = "Salary Estimate",
                 "company" = "Company Name", "location" = "Location",
                 "size" = "Size"
                 )
)

```


```{r}
# clean/merge data

jobs = dplyr::union(jobs_1, jobs_2)

jobs <- jobs |>
  separate_wider_delim(
    location, 
    names = c("city", "state_id"), 
    delim = ",",
    too_few = ("align_start"),
    too_many = ("drop")
    ) |>
  mutate(salary_est = (str_remove_all(salary_est, "[^0-9-K]"))) |>
  separate_wider_delim(
    salary_est,
    names = c("min", "max"),
    delim = "-",
    too_few = ("align_start"),
    too_many = ("drop")
  ) |>
  mutate(
    
    #no K means it is an hourly wage, which can closely approximate
    #annual wages by multiplying by 2000x. Probably better than NA
    
    min = case_when(
      is.na(min) ~ NA,
      str_detect(min, "K") ~ 1000*as.integer(str_remove_all(min, "[^0-9]")),
      .default = 2000*as.integer(min)
    ),
    max = case_when(
      is.na(max) ~ NA,
      str_detect(max, "K") ~ 1000*as.integer(str_remove_all(max, "[^0-9]")),
      .default = 2000*as.integer(max)
    ),
    avg = as.integer((min + max) / 2)
  ) |>
  relocate(avg, .after = max)


jobs <- jobs |>
  mutate(
    city = str_trim(city),
    state_id = str_trim(state_id)
  )

cities <- cities |>
  mutate(
    city = str_trim(city),
    state_id = str_trim(state_id)
  )


jobs <- jobs |>
  left_join(cities, join_by(city, state_id))
```

```{r}
jobs |>
  summarize(
    avg_salary_us = mean(avg, na.rm = TRUE)
  )
```
```{r}
jobs |> 
  count(city, name = "jobs") |>
  arrange(desc(jobs))
```



```{r}

#I think large means that it is a more granular representation

# Get US states
usa <- ne_states(country = "united states of america", returnclass = 'sf')
```



```{r}
jobs_plotting <- jobs |>
  group_by(city, state_id, lat, long) |>
  summarize(
    city_avg_salary = round(mean(avg, na.rm = TRUE), -2),
    city_num_jobs = n()
  ) |> 
  select(city, lat, long, city_avg_salary, city_num_jobs) |>
  filter(is.na(lat) == FALSE & is.na(long) == FALSE) |>
  arrange(desc(city_num_jobs))      #This is so leaflet puts the little 
                                    #circles on top of the bigger circles
  
```

```{r}
color_palette <- colorNumeric(
  palette = "inferno",
  domain = jobs_plotting$city_avg_salary    #finds range of avg_salary column
)


leaflet(data = jobs_plotting) %>%
  addTiles() %>%
  addCircles(
    ~long, 
    ~lat,
    popup = ~paste0(
      jobs_plotting$city, ", ", jobs_plotting$state_id, "<br/>",
      "Jobs: ", jobs_plotting$city_num_jobs, "<br/>",
      "Average salary: ", jobs_plotting$city_avg_salary
      ),
    radius = (jobs_plotting$city_num_jobs + 2)*250,
    opacity = 0.9,
    fillColor = ~color_palette(jobs_plotting$city_avg_salary),
    color = ~color_palette(jobs_plotting$city_avg_salary),
    weight = 2
    ) %>%
  setView(lng = -95, lat = 37, zoom = 4) %>%
  addLegend(
    pal = color_palette,            
    values = ~jobs_plotting$city_avg_salary, 
    opacity = 0.7,
    title = "Average salary",
    position = "bottomright"
  ) %>%
  addScaleBar()
```

```{r}
#Work in progress

#| warning: false

ggplot(data = usa) +
  geom_sf() +
  geom_text(data= jobs, aes(x=long, y=lat, label=city),
            color = "darkblue", fontface = "bold", check_overlap = FALSE) +
  coord_sf(xlim = c(-125, -66.6), ylim = c(24.2, 49.8), expand = FALSE)

```









